Entropy Rain: Dilution and Compression of Thermals in Stratified Domains

### 2019-07-18 submission2/referee_response.txt

"""
8. In the caption to Fig.4, I'd add "(AN)" and "(FC)" after anelastic
and fully compressible. The differences are indeed quite small. Maybe
it would be good to emphasize that the author mean 2D Cartesian and
not 2D cylindrical, if that is indeed the case. 2D Cartesian implies a
very different width of vortices, so it is amazing that this does not
matter. Given that AN and FC are changed at the same time as 2D and 3D,
it is not clear what is caused by what. I agree, the differences are
small, but a comment might be useful. Is 3-D AN harder, or what?
"""

We have added (AN) and (FC) to the captions of figures 4 and 5.
Furthermore, we have added some additional text to 3.1, as it seems that it was unclear that the 3D simulations are in azimuthally symmetric cylindrical domains, while the 3D simulations are in cartesian domains.

"""
9. Except for general remarks in the introduction, nothing is said about
the enthalpy flux that is accomplished by one plume. It can obviously
not be modeled by a superadiabatic gradient, but maybe some quantitative
statements can be made regarding some of the assumptions made in B16.
"""

We have added two panels to Figure 6 which explore the enthalpy flux achieved by a thermal, and the filling factor required to carry the solar luminosity.
An accompanying discussion has been added at the end of section 5.

### 2019-07-18 tex/stratified_thermals.tex
close to ready for resubmission

Thermals in the Boussinesq limit have been well studied in the laboratory for decades \citep[see e.g.][]{morton&all1956, scorer1957}, and more recently through Direct Numerical Simulation \citep[DNS,][]{lecoanet&jeevanjee2018}. 
These studies find that thermals expand radially and decelerate as they propagate.
Such a deceleration may cause the thermals to move so slowly they would diffuse away their entropy perturbation before reaching the bottom of the solar convection zone.
However, this past work does not include the effects of density stratification.

@@

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{thermal_sun_comparison.pdf}
    \caption{
	(top) Bounds on the evolution of the radii of entropy rain in the Sun are shown and compared to what the radii would be under purely horizontal compression.
	(middle) The corresponding range of velocities of solar-like thermals are shown. 
	(bottom) Shown is the estimated range of the ratio of the thermal diffusion timescale normalized by the freefall timescale over the thermal radial length scale.
	Diffusivities are calculated using the realistic solar-like diffusion profiles of \citet{brown2011}.
    \label{fig:thermal_sun_comparison} }
\end{figure}

Because solar downflows likely break up into thermals of various sizes, we will examine the fate of thermals with dimensionless buoyancy in the interval $[B_\ell, B_u] = 0.5B_\text{th}, 2B_\text{th}$.
Interpolating and extrapolating the data in Table \ref{table:parameters}, we use Eqns.~\ref{eqn:r_theory} \& \ref{eqn:theory_z} to calculate theoretical predictions for how thermals with these buoyancies would evolve over extended atmospheres like the solar convection zone.
Using a simple solar interior model calculated using MESA \citep{paxton&all2011} to map the density profiles of our simulation domains onto the density profile of the Sun, we plot the evolution of thermals in our estimated interval in Fig. \ref{fig:thermal_sun_comparison}.
In the top and middle panels of Fig. \ref{fig:thermal_sun_comparison}, we show the evolution of these thermals' radii and velocities inside of the Sun, and compare their evolution to pure horizontal compression.
We use the solar diffusivity models of \citet{brown2011} to estimate the timescale over which the thermal would diffuse its entropy signature ($\tau_\kappa = \chi/r_{\text{th}}^2$) and the thermal freefall timescale over its own radial length scale ($\tau_{\text{ff}} = r_{\text{th}}/w_{\text{th}}$).
In the bottom panel of Fig. \ref{fig:thermal_sun_comparison}, we plot the ratio of these two timescales over the thermal's evolution.

We find that our estimated solar thermal is in the falling regime.
These thermals experience radial compression with corresponding increases in velocity, but experience much less compression than a naive approximation based on the density stratification alone.
\textbf{Interestingly, the thermal's Mach number is roughly a value of 0.1 throughout the full extent of the solar convection zone.}
Furthermore, throughout the thermal's fall, we find that $\tau_\kappa \gg \tau_{\text{ff}}$, and thus we do not expect the thermal to diffuse its entropic signature.
We find a fractional diffusion rate over the thermal's transit of the solar convection zone of 
$$
\int_{R_\odot}^{0.7 R_\odot} \tau_\kappa^{-1} \frac{dr}{w_{th}} = [6.55, 1.57]\times 10^{-3}\,\text{for}\,[B_\ell, B_u].
$$
Thus, the thermal loses less than 1\% of its entropic signature to diffusion.
These estimates suggest that entropy rain could transport the solar luminosity.

@@

$\,$\newline$\,$
\begin{acknowledgements}
We'd like to thank Brett McKim and Nadir Jeevanjee for discussions about thermals and the laminar entrainment picture.
\textbf{We'd like to further thank Nadir Jeevanjee, Jeremy Goodman, and the anonymous referee whose careful comments greatly improved the scientific quality and clarity of this manuscript.}
This work was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program -- Grant 80NSSC18K1199.
This work was additionally supported by NASA LWS grant NNX16AC92G and by the National Science Foundation under grant No.~1616538. 
DL is supported by Princeton Center for Theoretical Science and Lyman Spitzer Jr.~postdoctoral fellowships.
Computations were conducted with support by the NASA High End Computing (HEC) Program through the NASA  Advanced Supercomputing (NAS) Division at Ames Research Center on Pleiades with allocation GID s1647.
\end{acknowledgements}

>>>>

Thermals in the Boussinesq limit have been well studied in the laboratory for decades \citep[see e.g.][]{morton&all1956, scorer1957}, and more recently through Direct Numerical Simulation \citep[DNS,][]{lecoanet&jeevanjee2018}. 
These studies find that thermals expand radially and decelerate as they propagate.
Such a deceleration may cause the thermals to move so slowly they would diffuse away their entropy perturbation before reaching the bottom of the solar convection zone.
\textbf{\citet{brandenburg&hazlehurst2001} found that hot, buoyant thermals in stratified domains behave qualitatively similar to Boussinesq thermals.
However, we are not aware of past work which carefully examines the effects of stratification on \emph{negatively} buoyant thermals.
}

@@

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{thermal_sun_comparison.pdf}
    \caption{
	(First panel) Bounds on the evolution of the radii of entropy rain in the Sun are shown and compared to what the radii would be under purely horizontal compression.
	(Second panel) The corresponding range of velocities of solar-like thermals are shown. 
	(Third panel) Shown is the estimated range of the ratio of the thermal diffusion timescale normalized by the freefall timescale over the thermal radial length scale.
	Diffusivities are calculated using the realistic solar-like diffusion profiles of \citet{brown2011}.
	\textbf{
	(Fourth panel) An estimate of the luminosity of these thermals, normalized by the solar luminosity, is shown.
	(Fifth panel) An estimate of the filling factor of thermals required to carry the solar luminosity.}
    \label{fig:thermal_sun_comparison} }
\end{figure}

Because solar downflows likely break up into thermals of various sizes, we will examine the fate of thermals with dimensionless buoyancy in the interval $[B_\ell, B_u] = 0.5B_\text{th}, 2B_\text{th}$.
Interpolating and extrapolating the data in Table \ref{table:parameters}, we use Eqns.~\ref{eqn:r_theory} \& \ref{eqn:theory_z} to calculate theoretical predictions for how thermals with these buoyancies would evolve over extended atmospheres like the solar convection zone.
Using a simple solar interior model calculated using MESA \citep{paxton&all2011} to map the density profiles of our simulation domains onto the density profile of the Sun, we plot the evolution of thermals in our estimated interval in Fig. \ref{fig:thermal_sun_comparison}.
In the first and second panels of Fig. \ref{fig:thermal_sun_comparison}, we show the evolution of these thermals' radii and velocities inside of the Sun, and compare their evolution to pure horizontal compression.
We use the solar diffusivity models of \citet{brown2011} to estimate the timescale over which the thermal would diffuse its entropy signature ($\tau_\kappa = \chi/r_{\text{th}}^2$) and the thermal freefall timescale over its own radial length scale ($\tau_{\text{ff}} = r_{\text{th}}/w_{\text{th}}$).
In the third panel of Fig. \ref{fig:thermal_sun_comparison}, we plot the ratio of these two timescales over the thermal's evolution.

We find that our estimated solar thermal is in the falling regime.
These thermals experience radial compression with corresponding increases in velocity, but experience much less compression than a naive approximation based on the density stratification alone.
\textbf{Interestingly, the thermal's Mach number is roughly a value of 0.1 throughout the full extent of the solar convection zone.}
Furthermore, throughout the thermal's fall, we find that $\tau_\kappa \gg \tau_{\text{ff}}$, and thus we do not expect the thermal to diffuse its entropic signature.
We find a fractional diffusion rate over the thermal's transit of the solar convection zone of 
$$
\int_{R_\odot}^{0.7 R_\odot} \tau_\kappa^{-1} \frac{dr}{w_{th}} = [6.55, 1.57]\times 10^{-3}\,\text{for}\,[B_\ell, B_u].
$$
Thus, the thermal loses less than 1\% of its entropic signature to diffusion.

\textbf{
We now briefly estimate the enthalpy flux carried by a thermal.
Following \citet{brandenburg2016}, we estimate the enthalpy flux of the thermal as $F_{\text{enth}} = \rho_0(z_th) T_0(z_th) w_{\text{th}} S_{1,\text{th}}$.
We estimate the thermal's entropic signature as $S_{1, th} = B_{\text{th}}/(\rho_0(z_{\text{th}}) V_{\text{th}} g/c_P)$, where the thermal volume is $V_{\text{th}} = m r_{\text{th}}^3$.
We take $\rho_0(z_{\text{th}})$, $T_0(z_{\text{th}})$, and $g/c_P$ from our solar MESA model at the depth of the thermal over its evolution.
We take $r_{\text{th}}$ and $w_{\text{th}}$ to be the thermal radius and velocity shown in the first and second panels of Fig.~\ref{fig:thermal_sun_comparison}.
We find $m$ from the data in Table \ref{table:parameters} in the same manner as we found $B_{\text{th}}$ for these solar thermals.
To estimate the total luminosity carried by one thermal, we calculate $L_{\text{enth, th}} = \pi r_{\text{th}}^2 F_{\text{enth}}$, and we plot this value as a function of depth, normalized by the solar luminosity, in panel 4 of Fig.~\ref{fig:thermal_sun_comparison}.
}

\textbf{
From our estimate of the luminosity carried by one thermal, we estimate the filling factor of downward propagating thermals required to carry the solar luminosity through their enthalpy flux.
We calculate $f_{\text{th}} = (L_{\odot}/L_{\text{enth, th}}) (\pi r_{\text{th}}^2)/(4\pi R^2)$, where $R$ is the thermal's radial distance form the center of the Sun.
We plot the $f_{\text{th}}$ in panel 5 of Fig.~\ref{fig:thermal_sun_comparison}.
The filling factor of thermals required to carry $L_\odot$ at the solar surface is greater than unity; this is unsurprising, as we know that solar surface convection carries the solar luminosity through the combined effects of upflows and downflows.
A few percentage of the solar radius beneath the photosphere, $f_{th}$ drops to a very modest $10^{-4}$, and by the base of the solar convection zone approaches $10^{-7}$.
These estimates suggest that even if a large fraction of the thermals launched from the solar surface break apart due to turbulence, this mechanism could still efficiently carry the solar luminosity deep in the convection zone.
}

\textbf{
While these estimates suggest that entropy rain could effectively transport the solar luminosity, we will briefly note that our handling of the enthalpy flux here ignores the contributions of the kinetic energy flux, potential energy flux, and viscous flux.
While we expect the last of these to be inconsequential, the kinetic energy flux (which transports luminosity inwards) and the potential energy flux (which transports luminosity outwards) may be large.
A more thorough examination of these fluxes are outside of the scope of this work, but due to the promising nature of our results found here for the entropy rain hypothesis, such work seems promising.
}

@@

$\,$\newline$\,$
\begin{acknowledgements}
We'd like to thank Brett McKim and Nadir Jeevanjee for discussions about thermals and the laminar entrainment picture.
\textbf{We'd like to further thank Nadir Jeevanjee, Axel Brandenburg, Jeremy Goodman, and the anonymous referee whose careful comments greatly improved the scientific quality and clarity of this manuscript.}
This work was supported by NASA Headquarters under the NASA Earth and Space Science Fellowship Program -- Grant 80NSSC18K1199.
This work was additionally supported by NASA LWS grant NNX16AC92G and by the National Science Foundation under grant No.~1616538. 
DL is supported by Princeton Center for Theoretical Science and Lyman Spitzer Jr.~postdoctoral fellowships.
Computations were conducted with support by the NASA High End Computing (HEC) Program through the NASA  Advanced Supercomputing (NAS) Division at Ames Research Center on Pleiades with allocation GID s1647.
\end{acknowledgements}
